Weakly Nonlinear MRI: Notes

### 2014-10-18 notes/mri_notes.tex
more hmri derivation
final momentum terms for hmri
induction eqn hmri

@@

>>>>

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

@@

The Lorentz term in the momentum equation is the only term that will change. It becomes

\beq
\left(\nabla \times \left(B_0 \hat{z} + \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} + \mathbf{B}_1
\right) \right) \times \left(B_0 \hat{z} + \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} + \mathbf{B}_1
\right)
\eeq

and we gain a potential 5 terms that were not present before, since there are 5 terms in this expansion that involve the azimuthal term. These are as follows.

\beq
\left(\nabla \times B_0 \hat{z} \right) \times \xi B_0 \left(\frac{x_1}{x}\right)\hat{y} = 0
\eeq

because of the constant term.

\beq
\left(\nabla \times \xi B_0 \left(\frac{x_1}{x}\right) \hat{y} \right) \times B_0 \hat{z} = \left(\partial_x \left(\xi B_0 \left(\frac{x_1}{x}\right)\right)\hat{z}\right) \times B_0 \hat{z} = 0
\eeq

because $\hat{z} \times \hat{z} = 0$.

\beq
\left(\nabla \times \xi B_0 \frac{x_1}{x} \hat{y} \right) \times \left( \xi B_0 \frac{x_1}{x}\hat{y}\right) = \xi^2 B_0^2 \frac{x_1^2}{x^3} \hat{x}
\eeq

And

\beq
\left(\nabla \times \xi B_0 \frac{x_1}{x} \hat{y}\right) \times \mathbf{B_1} = -\xi B_0 \frac{x_1}{x^2} \left(B_{1x} \hat{y} - B_{1y} \hat{x}\right)
\eeq

And finally,

\beq
\left(\nabla \times \mathbf{B}_1 \right) \times \left(\xi B_0 \frac{x_1}{x} \hat{y}\right) = - \partial_z B_{1y} \xi B_0 \frac{x_1}{x} \hat{z} - \partial_x B_{1y} \xi B_0 \frac{x_1}{x} \hat{x}
\eeq

So our final added terms to the momentum equation are:

\beq
\xi^2 B_0^2 \frac{x_1^2}{x^3} \hat{x} \, - \,  \xi B_0 \frac{x_1}{x^2} B_{1x} \hat{y} \, + \, \xi B_0 \frac{x_1}{x^2} B_{1y} \hat{x} \, - \, \partial_z B_{1y} \xi B_0 \frac{x_1}{x} \hat{z} \, - \, \partial_x B_{1y} \xi B_0 \frac{x_1}{x} \hat{x}
\eeq

For the induction equation, we have:

\beq
\partial_t \mathbf{B} = \left(\mathbf{B} \cdot \nabla \right) \mathbf{u} - \left(\mathbf{u} \cdot \nabla \right) \mathbf{B} + \frac{1}{\reym} \nabla^2 \mathbf{B}
\eeq

The first term on the righthand side becomes:

\beq
\partial_t \mathbf{B} = \left( \left( \mathbf{B_0} \hat{z} + \xi B_0 \frac{x_1}{x} \hat{y} + \mathbf{B_1} \right) \cdot \nabla \right) \left(-q \Omega_0 x \hat{y} + \mathbf{u_1}\right) = \xi B_0 \frac{x_1}{x} \partial_y \mathbf{u_1}
\eeq

The second term becomes:

\beq
\xi B_0 u_{1x} \frac{x_1}{x^2} \hat{y}
\eeq
